Update of the ozone dry deposition in the OsloCTM3 v1.0

### 2018-11-22 DryDep.tex
Almost finished comparison with mmm. Updated plot again.

For the temperature dependent dynamic viscosity of air $\mu(T)$, we chose a linear fit to Sutherland's law through the origin within the temperature range $\{T \in \mathbb{R} | (243.15 < T < 313.15)\,\unit{K}\}$: $\mu(T) = 6.2\cdot 10^{-8}\,\unit{kg\,m^{-1}\,s^{-1}\,K^{-1}} \cdot T$. But despite its rough accuracy, we found that the choice of $\mu(T)$ has no effect on $\left<R_b\right>$ (Supplement S.1). In Eq.~(\ref{eq:kinematic_viscosity}), $\rho$ is substituted by the air density using the ideal gas law. $P_0$ is the surface pressure, as $T$ the $2\,\unit{m}$ temperature is chosen, and $R_\text{air}$ is the universal gas constant for air.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Comparison with modeling results}
\label{subsec:model}
%
In our evaluation of results, we closly follow suggestions by \citet{ACP:Hardacre2015}. For the purpose of comparison with the multi-model mean of the therein participating Task Force on Hemispheric Transport of Air Pollution (TF HTAP) models, we also have re-grided our data to a horizontal resolution of $3^\circ\times 3^\circ$. In Section~\ref{subsubsec:zonal}, we look at zonal ditributions of $[\chem{O_3}](z_0)$, $v^\chem{O_3}_\text{DD}$, and $\sum\chem{O_3^\text{DD}}$ for all our senstivity simulations and study seasonal cycles of hemispheric ozone as well as nine land surface types which have been retrospectively seperated (Section~\ref{subsubsec:seasons}). From this, we estimate the total annual ozone dry deposition onto ocean, ice, and land surfaces and compare with results from \citet{ACP:Luhar2017}.
%
\subsubsection{Zonal distribution}
\label{subsubsec:zonal}
%
The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a) displays, in consistancy with Fig.~\ref{fig:diff_maps}a), a general, global increase of ozone by up to $10\,\unit{ppb}$ which is largest between $(25-50)\,\unit{^\circ N}$ including the major deserts. In the deep tropics ($5\,\unit{^\circ S}-5\,\unit{^\circ N}$), the increase is smallest ($\Delta[\chem{O_3}]\leq 5\,\unit{ppb}$). We find that the EMEP scheme further intensifies the strong asymmetry between northern and southern hemisphere as a consequence of the distribution of the continental land masses and vegetation thereon. Among the sensitivity studies, there is only a low variance. Most remarkable and as expected due to the strong decrease in prescribed dry deposition velocity over ice and snow, \emph{EMEP\_ppgssh\_ice} displays almost a doubling of surface ozone in the high Arctics compared to \emph{Wesely\_type} but affects ozone concentrations down to latitudes at about $50\,\unit{^\circ}$ in both hemispheres. Neglecting any dependence on solar radiation in the stomatal conductance formulation (\emph{EMEP\_offLight}) -- or in other words allowing photosynthesis 24/7 -- generally decreases the ozone concentration by about $2\,\unit{ppb}$, while choosing soil water at deeper levels than the surface layer (\emph{EMEP\_SWVL4}) increases $[\chem{O_3}](z_0)$ slightly. Rather surprisingly, switching off the phenology completely (\emph{EMEP\_offPhen}), doesn't make much difference. The scenario of differing emissions (2005 in comparison to 2014 or more specifically \emph{EMEP\_ppgs\_2005} compared to \emph{EMEP\_ppgs}), yields higher ozone concentrations in the northern hemisphere in 2005 in accordance to a reduction in sulfur and \chem{NO_x} emissions in south east Asia in later years. An opposite tendency is seen for latitudes south of $30\,\unit{^\circ N}$, propbably owing to the industrial development of countries in these regions.

Average dry deposition velocities have been reconstructed from the ratio between surface concentration and total dry deposition
\begin{equation}
  v^\chem{O_3}_\text{DD} = \frac{\sum \chem{O_3^\text{DD}}(z_0)}{\chem{O_3}(z_0)}\cdot c_\text{month},
\end{equation}
with $c_\text{month} = \frac{\text{layer thickness}}{\text{seconds in month}}$. This way of computing dry deposition velocities has certain drawbacks of which one is the freedom of choice of the \emph{layer thickness}. We chose the average thickness of the lowermost model level ($16\cdot10^{2}\,\unit{cm}$) to match the prescribed dry deposition velocities above ice and snow ($v_\text{ice/snow}^\chem{O_3} = 0.05\,\unit{cm\,s^{-1}}$). The resulting $v^\chem{O_3}_\text{DD}$ are shown in Fig.~\ref{fig:mmm_drydep}b). The dry deposition velocities in the EMEP scheme are much smaller than in the Wesely scheme for latitudes between $50\,\unit{^\circ S}$ and $70\,\unit{^\circ N}$. In the Arctics, except for \emph{EMEP\_ppgssh\_ice}, all realisations of the EMEP scheme are constantly above the multi-model-mean. This indicates, that with respect to other models, the updated dry deposition velocity above ice and snow should be considered as new standard for the OsloCTM3. The dry deposition velocities are of cause independent of the emission scenario, but display a strong dependence on $f_\text{light}$ and $f_\text{phen}$. The closer to $1$ these factores are the higher the dry deposition velocities become (compare \emph{EMEP\_offLight}, \emph{EMEP\_offPhen}, and \emph{EMEP\_ppgssh}). That highlights the importance of a proper PAR from meteorological input in CTMs or radiative transfer in PAR in global climate models (GCM) for canopy conductance computation. Albeit, the total numbers are smaller, the shape of the zonal average dry deposition velocities from the EMEP scheme and the multi-model-mean are similar (for surface tupe partitioning see figure in Supplement~S.6.1). The biggest exception is the zonal band $(12-30)\,\unit{^\circ S}$, which coincide with the location of Australia and its large desert regions, and its counterpart in the northern hemisphere.

The annual total ozone dry deposition is shown in Fig.~\ref{fig:mmm_drydep}b). In accordance to the priviously decribed features, we observe a reduction of the global total ozone dry deposition to almost half of \emph{Wesely\_type} in all sensitivity studies. The occurence of this reduction in the zonal bands where the major deserts are located may point to a important change in $v_\text{desert}^\chem{O_3}$. Consulting the Wsely scheme parameter file, we find $v_\text{desert}^\chem{O_3} \equiv v_\text{tundra}^\chem{O_3} = 0.26\,\unit{cm\,s^{-1}}$ \citep{JGR:Hough1991}, while in the EMEP scheme $v_\text{desert}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$ and $v^\chem{O_3}_\text{tundra} = 0.24\,\unit{cm\,s^{-1}}$, respectively. Similarly, dry deposition velocities over ice and snow and ocean have been even higher in the Wesely scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.07\,\unit{cm\,s^{-1}}$) than in the original EMEP scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$). These differences in baseline dry deposition velocities over huge parts of the unvegetated surface of the Earth (see Fig.~\ref{fig:pft_landsurface}) amount to most of the quantitative difference between the Wesely and the EMEP scheme. We further elaborate on this in the following section.

\begin{figure}[t]
  \includegraphics[width=8.3cm]{pictures/final-vmr_surfaceozone}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydep}
  \caption{Comparison of the manifold of OsloCTM3 integrations with respect to (a) Surface ozone concentrations, (b) Annual average ozone dry deposition velocity, (c) Total annual ozone dry deposition. The different colors indicate sets of simulation with similar baselines. The multi-model mean from the evaluation of HTAP models by \citet{ACP:Hardacre2015} is shown as a reference (where available).}
  \label{fig:mmm_drydep}
\end{figure}
%
In Fig.~\ref{fig:mmm_drydep_season}a), the average zonal ozone dry deposition is shown seperated by month. Where available, we have added the multi-model-mean given by \citet{ACP:Hardacre2015} as reference. There are two major phases apparent, NH and SH greening season. Former is reflected in a pronounced peak of $v_\text{DD}^\chem{O_3}$ in the northern mid latitudes, while the absence of this peak can be attributed to the latter in addition to a southward shift of the tropical peak dry deposition velocity. In the Wesely scheme, NH midlatitude peak velocities appear in June compared to July in the EMEP scheme, indicating that the prescribed seasonal cycles differ. The corresponding total monthly ozone dry deposition is shown in Fig.~\ref{fig:mmm_drydep_season}b). In general, the seasonal patterns are quite similar in the Wesely scheme and the EMEP scheme, displaying a strong symmetry around $10\,\unit{^\circ N}$ in January/February and November/December, respectively. What differs most is the molding and intensitity of the NH peak dry deposition. Both schemes reach the maximum in June/July but the peak is much more differenciated in March already in the Wesely scheme. Similarly, the SH tropical peak dry deposition is reached in August/September but sustained longer, into October, in the Wesely scheme. Since we have not conducted any simulation with a meteorological year other than 2005, we cannot elaborate on whether this is a special feature of our chosen year or not.
%
\begin{figure}[t]
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo_monthly}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydep_monthly}
  \caption{Comparison of the manifold of OsloCTM3 integrations with respect to (a) Zonal average ozone dry deposition velocities; (b) Total annual amount of ozone removed from the atmosphere via dry deposition. The multi-model mean from the evaluation of HTAP models by \citet{ACP:Hardacre2015} is shown as a reference (where available).}
  \label{fig:mmm_drydep_season}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Average seasonal cycles}
\label{subsubsec:seasons}
%
To further disentangle the contributions of different regions to the global ozone budget, we will briefly look at different projections of seasonal cycles.

In Fig.~\ref{fig:mmm_drydep_hem}, the total annual ozone dry deposition seperated into mid and high latitudes in the northern hemisphere ($(30-90)\,\unit{^\circ N}$), the tropics and subtropics ($30\,\unit{^\circ S}-30\,\unit{^\circ N}$), and the mid and high latitudes in the southern hemisphere ($(30-90)\,\unit{^\circ S}$) are shown. As reference, we have added the multi-model-mean by \citet{ACP:Hardacre2015}. As expected, the NH mid and high latitudes display a strongly pronounced seasonal cycle, while it is less pronounced in the tropics (due to the lack of seasons) and in the SH (due to the small percentage of vegetated surface). The highest ozone dry deposition is found in the tropics and amounts on average to the  peak level of dry deposition in the NH for the multi-model-mean \citep{ACP:Hardacre2015} and EMEP scheme. In the Wesely scheme, the average tropical ozone dry deposition diverges by about $5\,\unit{Tg}$ in comparison to its corresponding NH maximum. The seasonal cycle in the EMEP scheme appears to be shifted towards later in the year in the NH. While the general seasonal cycle in the tropics and subtropics is not changing, we find the strongest sensitivity to the various realisations of the EMEP scheme there, with \emph{EMEP\_offLight} and \emph{EMEP\_ppgs\_2005} the highest and the lowest, respectively. This indicates that surface ozone in the tropics may be very sensitive to changes to emissions and changes in longwave radiative transfer (e.g., cloud cover, aerosol).

%
\begin{figure*}[t]
  \includegraphics[width=12cm]{pictures/final-total_ozone_drydep_hem}
  \caption{Seasonal cycle of total annual amount of ozone removed from the atmosphere through dry deposition separated into northern hemisphere (NH), tropics (TR), and southern hemisphere (SH). The multi-model mean from the evaluation of HTAP models by \citet{ACP:Hardacre2015} is shown as a reference.}
  \label{fig:mmm_drydep_hem}
\end{figure*}
%
As suggested by \citet{ACP:Hardacre2015}, we retrospectively separate ozone dry deposition velocities with respect to surface types. Based on a CLM~2 average dynamic land surface map, we generated masks for 9 different surface types (Fig.~\ref{fig:pft_landsurface}a) and used these to select gridboxes with a high percentage of these surface types, ranging from a meager $22\,\unit{\%}$ for tundra in the southern hemisphere to $98\,\unit{\%}$ regarding ocean. From this CLM~2 based map, it was not possible to select gridboxes with $100\,\unit{\%}$ for each surface type. Since we have not perform a full unfolding on the data, the results should be treated with slight caution. In Fig.~\ref{fig:mmm_drydep_season_pft}, the seasonal cycles of dry deposition velocities are shown for the nine surface categories. The patterns and absolute numbers differ between the schemes and the multi-model-mean. Exceptually large is the divergence for desert regions. The multi-model-mean displays a very pronounced seasonal cycle ($0.15\,\unit{cm\,s^{-1}} \leq v^\chem{O_3}_\text{desert}\leq 0.10\,\unit{cm\,s^{-1}}$). The Wesely scheme as well as the EMEP scheme do not indicate any seasonal cycle. The estimated dry deposition velocities over desert regions are consistent with the previously mentioned baseline values of ozone dry deposition,  about one order of magnitude lower than in the EMEP scheme than in the Wesely scheme. From measurements of ozone fluxes in the Sahara desert, $v^\chem{O_3}_\text{desert, day} = 0.1\,\unit{cm\,s^{-1}}$, $v^\chem{O_3}_\text{desert, night} = 0.04\,\unit{cm\,s^{-1}}$, and $\left<v^\chem{O_3}_\text{desert}\right> = 0.065\,\unit{cm\,s^{-1}}$ have been deduced \citep{AE:Gusten1995}. This implies, that ozone dry deposition over desert regions is highly overestimated in the Wesely scheme, while it may be underestimated in the EMEP scheme. Similarly, the dry deposition velocities over water differ. From measurements over the ocean a value of $v^\chem{O_3}_\text{water} = 0.019\,\unit{cm\,s^{-1}}$ has been deduced \citep{JGR:Helmig2012}. In a comparison of various dry deposition mechanisms over ocean waters implemented in their model, \citet{ACP:Luhar2017} give a range of $0.018\,\unit{cm\,s^{-1}} \leq \left<v^\chem{O_3}_\text{water}\right> \leq 0.039\,\unit{cm\,s^{-1}}$.
%
\begin{figure*}[t]
  \includegraphics[width=12cm]{pictures/final-total_ozone_drydepvelo_monthly_pft}
  \caption{Average ozone dry deposition velocities separated by land use type and compared to results from \citep{ACP:Hardacre2015}.}
  \label{fig:mmm_drydep_season_pft}
\end{figure*}
%

%
\begin{table*}[t]
\caption{Total ozone dry deposition for the respective year in \unit{Tg\,a^{-1}} separated into ocean, ice and, land contributions.}
\begin{tabular}{lcccccccccccc}
\tophline
\multirow{3}{*}{Simulation} & \multicolumn{3}{c}{Ocean} & \multicolumn{3}{c}{Ice} & \multicolumn{3}{c}{Land} & \multicolumn{3}{c}{Total}\\
& NH & SH & Global & NH & SH & Global & NH & SH & Global & NH & SH & Global\\
& \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})}\\
\middlehline
Wesely\_type & 88.7 & 110.3 & 199.0 & 0.7 & 5.0 & 5.7 & 346.1 & 153.5 & 499.6 & 612.4 & 347.9 & 960.2 \\
EMEP\_full & 67.1 & 84.7 & 151.8 & 0.6 & 4.4 & 5.0 & 236.0 & 112.3 & 348.2 & 408.6 & 245.6 & 654.2 \\
EMEP\_offLight & 66.6 & 84.1 & 150.7 & 0.6 & 4.4 & 5.0 & 267.8 & 129.8 & 397.6 & 451.3 & 267.2 & 718.6 \\
EMEP\_offPhen & 67.1 & 84.6 & 151.7 & 0.6 & 4.4 & 5.0 & 239.9 & 116.0 & 355.9 & 413.9 & 249.9 & 663.7 \\
EMEP\_SWVL4 & 68.1 & 86.1 & 154.2 & 0.6 & 4.5 & 5.1 & 241.8 & 115.0 & 356.8 & 417.2 & 250.7 & 667.9 \\
EMEP\_ppgs & 67.1 & 84.6 & 151.7 & 0.6 & 4.4 & 5.0 & 238.7 & 112.6 & 351.3 & 411.4 & 246.0 & 657.4 \\
EMEP\_ppgssh & 67.0 & 84.5 & 151.6 & 0.6 & 4.4 & 5.0 & 243.9 & 119.4 & 363.3 & 419.4 & 253.6 & 673.0 \\
EMEP\_ppgssh\_ice & 66.5 & 84.8 & 151.3 & 0.2 & 1.3 & 1.4 & 240.3 & 120.4 & 360.7 & 413.0 & 251.1 & 664.1 \\
EMEP\_ppgs\_2005 & 66.8 & 83.5 & 150.4 & 0.6 & 4.4 & 5.0 & 237.8 & 110.0 & 347.8 & 409.2 & 240.7 & 649.9 \\
\bottomhline
\end{tabular}
%\belowtable{} % Table Footnotes
\end{table*}
%

>>>>

For the temperature dependent dynamic viscosity of air $\mu(T)$, we chose a linear fit to Sutherland's law through the origin within the temperature range $\{T \in \mathbb{R} | (243.15 < T < 313.15)\,\unit{K}\}$: $\mu(T) = 6.2\cdot 10^{-8}\,\unit{kg\,m^{-1}\,s^{-1}\,K^{-1}} \cdot T$. But despite its rough accuracy, we found that the choice of $\mu(T)$ has no effect on $\overline{R}_b$ (Supplement S.1). In Eq.~(\ref{eq:kinematic_viscosity}), $\rho$ is substituted by the air density using the ideal gas law. $P_0$ is the surface pressure, as $T$ the $2\,\unit{m}$ temperature is chosen, and $R_\text{air}$ is the universal gas constant for air.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Comparison with modeling results}
\label{subsec:model}
%
In our evaluation of results, we closly follow suggestions by \citet{ACP:Hardacre2015}. For the purpose of comparison with the multi-model mean of the therein participating Task Force on Hemispheric Transport of Air Pollution (TF HTAP) models, we also have re-grided our data to a horizontal resolution of $3^\circ\times 3^\circ$. In Section~\ref{subsubsec:zonal}, we look at zonal ditributions of $[\chem{O_3}](z_0)$, $v^\chem{O_3}_\text{DD}$, and $\sum\chem{O_3^\text{DD}}$ for all our senstivity simulations and study seasonal cycles of hemispheric ozone as well as nine land surface types which have been retrospectively seperated (Section~\ref{subsubsec:seasons}). From this, we estimate the total annual ozone dry deposition onto ocean, ice, and land surfaces and compare with results from \citet{ACP:Luhar2017}.
%
\subsubsection{Zonal distribution}
\label{subsubsec:zonal}
%
The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a) displays, in consistancy with Fig.~\ref{fig:diff_maps}a), a general, global increase of ozone by up to $10\,\unit{ppb}$. This increase is largest in the zonal band $(25-50)\,\unit{^\circ N}$ which contains the major deserts. In the deep tropics ($5\,\unit{^\circ S}-5\,\unit{^\circ N}$), the increase is smallest ($\Delta[\chem{O_3}]\leq 5\,\unit{ppb}$). We find that the EMEP scheme further intensifies the strong asymmetry between northern and southern hemisphere as a consequence of the distribution of the continental land masses and vegetation thereon. Among the sensitivity studies, there is only a low variance. Most remarkable but expected due to the much smaller prescribed dry deposition velocity over ice and snow, \emph{EMEP\_ppgssh\_ice} displays almost a doubling of surface ozone in the high Arctics compared to \emph{Wesely\_type} but affects ozone concentrations down to latitudes at about $50\,\unit{^\circ}$ in both hemispheres. Neglecting any dependence on solar radiation in the stomatal conductance formulation (\emph{EMEP\_offLight}) -- or in other words allowing photosynthesis 24/7 -- generally decreases the ozone concentration by about $2\,\unit{ppb}$, while choosing soil water at deeper levels than the surface layer (\emph{EMEP\_SWVL4}) increases $[\chem{O_3}](z_0)$ slightly. Rather surprisingly, switching off the phenology completely (\emph{EMEP\_offPhen}), amounts only to small difference. The scenario of differing emissions (2005 in comparison to 2014 or more specifically \emph{EMEP\_ppgs\_2005} compared to \emph{EMEP\_ppgs}), yields higher ozone concentrations in the northern hemisphere in 2005 in accordance to a reduction in sulfur and \chem{NO_x} emissions in south east Asia in later years. An opposite tendency is seen for latitudes south of $30\,\unit{^\circ N}$, propbably owing to the industrial development of countries in these regions.

Average dry deposition velocities have been reconstructed from the ratio between surface concentration and total dry deposition
\begin{equation}
  v^\chem{O_3}_\text{DD} = \frac{\sum \chem{O_3^\text{DD}}(z_0)}{\chem{O_3}(z_0)}\cdot c_\text{month},
\end{equation}
with $c_\text{month} = \frac{\text{layer thickness}}{\text{seconds in month}}$. This way of retrospective computing of dry deposition velocities has certain drawbacks of which one is the freedom of choice of the \emph{layer thickness}. We chose the average thickness of the lowermost model level ($16\cdot10^{2}\,\unit{cm}$) to match the prescribed dry deposition velocities above ice and snow ($v_\text{ice/snow}^\chem{O_3} = 0.05\,\unit{cm\,s^{-1}}$). The resulting $v^\chem{O_3}_\text{DD}$ are shown in Fig.~\ref{fig:mmm_drydep}b). The dry deposition velocities in the EMEP scheme are much smaller than in the Wesely scheme for latitudes between $50\,\unit{^\circ S}$ and $70\,\unit{^\circ N}$. In the Arctics, except for \emph{EMEP\_ppgssh\_ice}, all realisations of the EMEP scheme are constantly above the multi-model-mean. This indicates, that with respect to other models, the updated dry deposition velocity above ice and snow should be considered as new standard for the OsloCTM3. The dry deposition velocities are of cause independent of the emission scenario, but display a strong dependence on $f_\text{light}$ and $f_\text{phen}$. The closer to $1$ these factores are the higher the dry deposition velocities become (compare \emph{EMEP\_offLight}, \emph{EMEP\_offPhen}, and \emph{EMEP\_ppgssh}). That highlights the importance of a proper PAR from meteorological input in CTMs or radiative transfer in PAR in global climate models (GCMs) for canopy conductance computation. Albeit, the total numbers are smaller, the shape of the normalized zonal average dry deposition velocities of the EMEP scheme and the multi-model-mean are similar (Supplement~S.6.1). The biggest exception is the zonal band $(12-30)\,\unit{^\circ S}$, which coincides with the location of Australia and its desert regions, as well as its counterpart in the northern hemisphere. However, it is not clear whether these roughly $10\,\unit{\%}$ of desert in the SH in comparison with $80\,\unit{\%}$ being covered by ocean in the same zonal band could have such an impact.

The annual total ozone dry deposition is shown in Fig.~\ref{fig:mmm_drydep}b). In accordance to the priviously decribed features, we observe a reduction of the global total ozone dry deposition in all sensitivity studies to almost one half of te amount given by \emph{Wesely\_type}. The occurence of this reduction in the zonal bands where the major deserts are located points to a change in $v_\text{desert}^\chem{O_3}$. Consulting the parameter file used in the Wesely scheme, we find $v_\text{desert}^\chem{O_3} \equiv v_\text{tundra}^\chem{O_3} = 0.26\,\unit{cm\,s^{-1}}$ \citep{JGR:Hough1991}, while in the EMEP scheme $v_\text{desert}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$ and $v^\chem{O_3}_\text{tundra} = 0.24\,\unit{cm\,s^{-1}}$, respectively. Similarly, dry deposition velocities over ice and snow and ocean have been even higher in the Wesely scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.07\,\unit{cm\,s^{-1}}$) than in the original EMEP scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$). These differences in baseline dry deposition velocities over huge parts of the unvegetated surface of the Earth amount to most of the quantitative difference between the Wesely and the EMEP schemebut may not sufficiently. We further elaborate on this in the following section.

\begin{figure}[t]
  \includegraphics[width=8.3cm]{pictures/final-vmr_surfaceozone}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydep}
  \caption{Comparison of the manifold of OsloCTM3 integrations with respect to (a) Surface ozone concentrations, (b) Annual average ozone dry deposition velocity, (c) Total annual ozone dry deposition. The different colors indicate sets of simulation with similar baselines. The multi-model mean from the evaluation of HTAP models by \citet{ACP:Hardacre2015} is shown as a reference (where available).}
  \label{fig:mmm_drydep}
\end{figure}
%
In Fig.~\ref{fig:mmm_drydep_season}a), the average zonal ozone dry deposition is shown seperated by month. Where available, we have added the multi-model-mean given by \citet{ACP:Hardacre2015} as reference. There are two major phases apparent, NH and SH greening season. Former is reflected in a pronounced peak of $v_\text{DD}^\chem{O_3}$ in the northern mid latitudes, while the absence of this peak can be attributed to the latter in addition to a southward shift of the tropical peak dry deposition velocity. In the Wesely scheme, NH mid latitude peak velocities appear in June compared to July in the EMEP scheme, indicating that the prescribed seasonal cycles differ. The corresponding total monthly ozone dry deposition is shown in Fig.~\ref{fig:mmm_drydep_season}b). In general, the seasonal patterns are quite similar in the Wesely scheme and the EMEP scheme, displaying a strong symmetry around $10\,\unit{^\circ N}$ in January/February and November/December, respectively. What differs most is the molding and intensitity of the NH peak dry deposition. Both schemes reach the maximum in June/July but the peak is much more differenciated in March already in the Wesely scheme. Similarly, the SH tropical peak dry deposition is reached in August/September but sustained longer, into October, in the Wesely scheme. Since we have not conducted any simulation with a meteorological year other than 2005, we cannot elaborate on whether this is a special feature of our chosen year or not. In comparison, the available two month of the multi-model-mean display a different pattern. Especially the NH/SH asymmetry is much more pronounced in February, than it is the EMEP scheme.
%
\begin{figure}[t]
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo_monthly}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydep_monthly}
  \caption{Comparison of the manifold of OsloCTM3 integrations with respect to (a) Zonal average ozone dry deposition velocities; (b) Total annual amount of ozone removed from the atmosphere via dry deposition. The multi-model mean from the evaluation of HTAP models by \citet{ACP:Hardacre2015} is shown as a reference (where available).}
  \label{fig:mmm_drydep_season}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Average seasonal cycles}
\label{subsubsec:seasons}
%
To further disentangle the contributions of different regions to the global ozone budget, we will briefly look at different projections of seasonal cycles.

In Fig.~\ref{fig:mmm_drydep_hem}, the total annual ozone dry deposition seperated into mid and high latitudes in the northern hemisphere ($(30-90)\,\unit{^\circ N}$), the tropics and subtropics ($30\,\unit{^\circ S}-30\,\unit{^\circ N}$), and the mid and high latitudes in the southern hemisphere ($(30-90)\,\unit{^\circ S}$) is shown. We have added the multi-model-mean by \citet{ACP:Hardacre2015} as reference. As expected, the NH mid and high latitudes display a strongly pronounced seasonal cycle, while it is less pronounced in the tropics (due to the lack of seasons) and in the SH (due to the small percentage of vegetated surface). The highest ozone dry deposition is found in the tropics and amounts on average to the  peak level of dry deposition in the NH for the multi-model-mean \citep{ACP:Hardacre2015} and EMEP scheme. In the Wesely scheme, the average tropical ozone dry deposition diverges by about $5\,\unit{Tg}$ in comparison to its corresponding NH maximum. The seasonal cycle in the EMEP scheme appears to be shifted towards later in the year in the NH. The seasonal cycle in the tropics and subtropics is not changing, but the total amount of dry deposition of ozone is very sensitive to differing realisations of the EMEP scheme, with \emph{EMEP\_offLight} and \emph{EMEP\_ppgs\_2005} displaying the highest and the lowest amount, respectively. This indicates that surface ozone in the tropics may be very sensitive to changes to emissions and changes in longwave radiative transfer (e.g., due changes in cloud cover or aerosol).
  \includegraphics[width=12cm]{pictures/final-total_ozone_drydep_hem}
  \caption{Seasonal cycle of total annual amount of ozone removed from the atmosphere through dry deposition separated into northern hemisphere (NH), tropics (TR), and southern hemisphere (SH). The multi-model mean from the evaluation of HTAP models by \citet{ACP:Hardacre2015} is shown as a reference.}
  \label{fig:mmm_drydep_hem}
As suggested by \citet{ACP:Hardacre2015}, we retrospectively separate ozone dry deposition velocities with respect to surface types. Based on a CLM~2 average dynamic land surface map, we generated masks for 9 different surface types (Fig.~\ref{fig:pft_landsurface}a) and used these to select gridboxes with a high percentage of these surface types, ranging from a meager $70\,\unit{\%}$ for cropland in the NH mid latitudes to $100\,\unit{\%}$ regarding desert, ocean, snow and ice, and tropical forest. From this CLM~2 based map, it was not possible to exclusively select gridboxes with $100\,\unit{\%}$ cover for every surface type. Since we have not performed a full unfolding on the data, the results should be treated with slight caution. In Fig.~\ref{fig:mmm_drydep_season_pft}, the seasonal cycles of dry deposition velocities are shown for the nine surface categories. The patterns and absolute numbers differ substantially between the Wesely scheme and the EMEP scheme and the multi-model-mean. In our sensitivity study, the divergence of the average dry deposition velocities in desert regions ($\Delta \overline{v}_\text{desert}^\chem{O_3} = 0.14\,\unit{cm\,s^{-1}}$) as well as grassland ($\Delta \overline{v}_\text{desert}^\chem{O_3} = 0.4\,\unit{cm\,s^{-1}}$) is quite remarkable as well as the difference to the multi-model-mean in tropical forest regions ($\Delta \overline{v}_\text{desert}^\chem{O_3} = 0.38\,\unit{cm\,s^{-1}}$). In contrast to the Wesely scheme and the EMEP scheme, the multi-model-mean indicates a very pronounced seasonal cycle in desert regions ($0.15\,\unit{cm\,s^{-1}} \leq v^\chem{O_3}_\text{desert}\leq 0.10\,\unit{cm\,s^{-1}}$). The estimated dry deposition velocities over desert regions are consistent with the previously mentioned baseline values of ozone dry deposition, which means that in the EMEP scheme they are about one order of magnitude lower than in the Wesely scheme. From measurements of ozone fluxes in the Sahara desert \citep{AE:Gusten1995}, $v^\chem{O_3}_\text{desert, day} = 0.1\,\unit{cm\,s^{-1}}$, $v^\chem{O_3}_\text{desert, night} = 0.04\,\unit{cm\,s^{-1}}$, and $\overline{v}^\chem{O_3}_\text{desert} = 0.065\,\unit{cm\,s^{-1}}$ have been deduced. This implies, that ozone dry deposition over desert regions is highly overestimated in the Wesely scheme as well as in the multi-model-mean, while it may be underestimated in the EMEP scheme. Similarly, the dry deposition velocities over water differ. From measurements over the ocean a mean value of $\overline{v}^\chem{O_3}_\text{water} = 0.019\,\unit{cm\,s^{-1}}$ has been deduced \citep{JGR:Helmig2012}. In a model study of dry deposition mechanisms to ocean waters by means of prescribed $v_\text{water}^\chem{O_3}$ and one- and two-layer gas exchange modeling, \citet{ACP:Luhar2017} found $\overline{v}_\text{water}^\chem{O_3}$ ranging between $0.018\,\unit{cm\,s^{-1}}$ (two-layer scheme) and $0.039\,\unit{cm\,s^{-1}}$ (prescribed). With $\overline{v}^\chem{O_3}_\text{water} = (0.038\pm0.002)\,\unit{cm\,s^{-1}}$, the EMEP scheme (Section~\ref{subsubsec:Rb}) yields probably a too strong dry deposition to ocean which implicates that ozone concentrations might even become larger and dry deposition even lower in the model if a two-layer scheme would be implemented. With respect to forests, we might be able to improve the model performance further by allowing for the full range of generic tree types of the EMEP scheme, instead of using a broad average. 
  \includegraphics[width=12cm]{pictures/final-total_ozone_drydepvelo_monthly_pft}
  \caption{Average seasonal cycles of ozone dry deposition velocities separated by land use type. Results from \citep{ACP:Hardacre2015} are shown as a reference. We refrain from showing the full extend of \emph{EMEP\_offLight} here, since it is an extreme scenario and has been discussed already.}
  \label{fig:mmm_drydep_season_pft}
